\mytitle{}

### 2019-12-12 paper/paper.tex
Fix URLs in online cites.

	\ifentrytype{mics}{}{% Remove publisher and editor except for books
		\clearfield{url}
	}
}

@@

@@

@@

\title{\mytitle{}}
\author{
%	Kyle~A.~Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
	Kyle~A.~Simpson,
	Simon~Rogers,
	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
	School of Computing Science, University of Glasgow, Glasgow, G12 8RZ, UK\\
%	University of Glasgow, Glasgow, Scotland,\\
	\email{k.simpson.1@research.gla.ac.uk}
%Anonymous Giraffe,
%Anonymous Badger,
%Anonymous Ibex\\
%Unnamed Department, Nowhere\\
%\email{giraffe.a@unnamed.com}
}

@@

%?? Discuss the most important conclusions before the outline.
We find that the existing work for this task \cite{DBLP:journals/eaai/MalialisK15} fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
\end{itemize}

@@

Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed to appear as though they originated from the victim, causing many large replies to be sent to the intended target---significantly increasing a botnet's throughput while masking the identity of each participant.
\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

\subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$, it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This model of system interaction is best described as a \emph{Markov Decision Process} (MDP).
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal if an agent needs to further explore some region of the state space.
The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new, custom-built training corpus.

@@

Computing the approximate value of every action available in the current state forms the basis of a policy.
Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might use some other mechanism.
\Cref{fig:tilecode-select} extends the prior working example to show how the value of each action is computed (and which action would be chosen by a greedy policy), combining a global estimate ($T_{\mathit{bias}}$) with knowledge particular to each state.

\begin{figure}
	\centering
	\resizebox{0.5\linewidth}{!}{
		\begin{tikzpicture}		
		% Top half
		\def\startacs{
			{0.8},
			{0.3},
			{-0.4}%
		}

@@

\subsection{Threat Model}
An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to any host which connects to it, and they are expected to act rationally in its pursuit.
Threat actors are external and act intentionally, aren't expected to be classified as an \emph{advanced persistent threat}, and likely range from hacktivists to moderately funded adversaries.
We assume that attacks will be volumetric DDoS attacks targeting bandwidth with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication/collaboration with \nth{3} party organisations such as upstream ISPs.
The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.

We do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.
These are APT-level capabilities, whose exploration presents a rich source for future work.

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

\subsection{System Design and Assumptions}
A deployment environment is defined as a network, with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
These destination nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
Agents are co-located with each egress switch, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, $U_s$.

We assume that the deployment environment is a moderately complex software-defined network, because the paradigm offers features which can directly benefit RL agents acting within.
The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners.
Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.

\subsection{Algorithm}
To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than making use of neural networks or more complicated function approximators.
Exploration is introduced by making use of $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their weight vector updates with one another (but may share experience and traces with one another).

Although the choice of a classical RL method likely brings lower theoretical performance, there are significant reasons to favour such methods; these include lower latency decision-making, lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.

@@

%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}) depending upon the bandwidth measurements regularly received from monitors in the environment.
For any flow, an agent then computes the path it would take through the network.
The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) a middle hop or the last hop.
\Cref{fig:global-state-path} illustrates the process.

@@

We describe and analyse each of the per-flow features included in the state vector throughout \cref{sec:rethinking-the-state-space}.
Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.

@@

\subsection{Action Space}
When monitoring a source-destination pair, an agent uses its state vector to decide which proportion of that flow's \emph{inbound} traffic should be dropped.
%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
This is implemented by installing an action via OpenFlow, instructing its host switch to drop each relevant packet with probability $p$.
We choose to drop packets rather than impose traffic limits as it offers us a discrete action space without prior knowledge of traffic characteristics or measurement.
Furthermore, we need not consider burstiness, fairness or implementation costs (such as per-flow bucket sizes) which could limit scalability.
We offer two models on how to choose $p$:

\subsubsection{Instant control}
Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
These choices primarily ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
Since this model requires no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely opportunistic/myopic).

\subsubsection{Guarded control}
The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$,.
The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.

We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.

@@

\subsection{Systems Considerations}
Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
This introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated.
At their worst, these risks present additional attack surface to an adversary.
To adapt to these problems, we make use of \emph{timed random sequential} updates.

Each agent begins with an empty work list.
For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
Uncompleted work is passed on to the next timestep, until the list is emptied, at which point is is repopulated using the set of available measurements.
To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
We expect this should be tuned based on the frequency at which statistics arrive.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-cap-box}
	\vspace{-1cm}
	\caption{
		Learned performance of Instant Control agents when benign traffic is TCP-like, using only a single feature as a basis for decisions.
		All of the chosen features can offer a marked improvement over no protection at all.
		Global state and Mean IAT still offer the greatest improvement above baseline, but packet-level statistics are considerably less effective for this class of traffic.
		\label{fig:tcp-cap-feature-plots}
	}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{../plots/ftprep-tcp-laf-cap-box}
	\vspace{-1cm}
	\caption{
		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken as a basis for decisions.
		This combination causes a significant improvement in the effectiveness of packet-level and per-window statistics.
		\label{fig:tcp-laf-feature-plots}
	}
\end{figure}

The main element required to move to a per-source model is a feature set which offers high predictive power, such that behavioural differences are readily apparent to an agent.
Elaborating further on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and offer an investigation into their use alongside different traffic types:
%?? We use these features, and why...

@@

\fakepara{$\mathbf{\Delta}$ Send/receive rate}
The change in traffic rates caused by the last action.
Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible in these fields.

@@

\fakepara{Mean packet size per window}
Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.

>>>>

	\ifentrytype{misc}{}{% Remove url except for misc and online.
		\ifentrytype{online}{}{
			\clearfield{url}
		}
	}
}

@@

\usepackage{scalerel}
\usetikzlibrary{svg.path}

\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
	orcidlogo/.pic={
		\fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
		\fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
		svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
		svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
	}
}

\newcommand\orcidicon[1]{
	\href{https://orcid.org/#1}{\mbox{\scalerel*{
		\begin{tikzpicture}[yscale=-1,transform shape]
		\pic{orcidlogo};
		\end{tikzpicture}
	}{|}}}}

@@

%\usepackage{framed}
%\usepackage{adjustbox}
%\colorlet{shadecolor}{uofgcobalt!20}
\usepackage[color]{changebar}
\setlength\changebarsep{10pt}
\setlength\changebarwidth{6pt}
%\cbcolor{uofgcobalt}
\cbcolor{white}
\colorlet{revisiontext}{black}
%\colorlet{revisiontext}{uofgcobalt}

@@

\title{\mytitle{}}
\author{
%	Kyle~A.~Simpson\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant number EP/M508056/1].},
	\protect Kyle~A.~Simpson\protect\orcidicon{0000-0001-8068-9909},
%	Kyle~A.~Simpson,
	Simon~Rogers\protect\orcidicon{0000-0003-3578-4477},
%	Simon~Rogers,
	Dimitrios~P.~Pezaros\protect\orcidicon{0000-0003-0939-378X},~\IEEEmembership{Senior~Member,~IEEE,},\\
%	Dimitrios~P.~Pezaros,~\IEEEmembership{Senior~Member,~IEEE,},\\
	School of Computing Science, University of Glasgow, Glasgow, G12 8RZ, UK\\
%	University of Glasgow, Glasgow, Scotland,\\
	\email{k.simpson.1@research.gla.ac.uk}
%Anonymous Giraffe,
%Anonymous Badger,
%Anonymous Ibex\\
%Unnamed Department, Nowhere\\
%\email{giraffe.a@unnamed.com}
}

@@

%?? Discuss the most important conclusions before the outline.
We find that the existing work for this task \cite{DBLP:journals/eaai/MalialisK15} fails to account for congestion-aware traffic (i.e., TCP) and environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
To remedy this, we make throttling decisions on a per-source basis and present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!

\subsection{Contributions}
\cbstart
{\color{revisiontext}
This paper contributes two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
These are designed to make effective decisions irrespective of protocol, and act on individual flows at the edge of any network topology.
We offer an in-depth investigation into suitable features for automatic DDoS mitigation, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
These features have been suggested by past studies, and independently tested in their own contexts.
Our study is the first attempt to quantify the individual efficacy of each in an RL setting.

We implement reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
To our knowledge, this is the first attempt to study or replicate Opus-based VoIP traffic, which has become commonplace since the codec's release in 2012.
These new traffic models inform an empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
We additionally compare our work against SPIFFY \cite{DBLP:conf/ndss/KangGS16}, reuniting two divergent strands of research and grounding the study of RL-based DDoS \cbend defences.}

@@

Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
\emph{Amplification attacks} exploit services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most exploitable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed to appear as though they originated from the victim, causing many large replies to be sent to the intended target---significantly increasing a botnet's throughput while masking the identity of each participant.
\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

\subsection{Reinforcement Learning}\label{sec:reinforcement-learning}
\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with training an agent to choose an optimal sequence of actions in pursuit of a given task \cite{RL2E}.
We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$, it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This model of system interaction is a \emph{Markov Decision Process} (MDP).
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal if an agent needs to further explore some region of the state space.
The policy is refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
In practice, this means that agents are able to automatically adapt to evolving problems without operator intervention or a new, custom-built training corpus.

@@

Computing the approximate value of every available action forms the basis of a policy.
Actions with maximal value can be chosen each time (the \emph{greedy} policy), we might modify this by taking random actions with probability $\epsilon$ to encourage early exploration (the \emph{$\epsilon$-greedy policy}), or we might use some other mechanism.
\Cref{fig:tilecode-select} extends the prior working example to show how the value of each action is computed (and which action would be chosen by a greedy policy), combining a global estimate ($T_{\mathit{bias}}$) with knowledge particular to each state.

\begin{figure}
	\centering
	\resizebox{0.45\linewidth}{!}{
		\begin{tikzpicture}		
		% Top half
		\def\startacs{
			{0.8},
			{0.3},
			{-0.4}%
		}

@@

\subsection{Threat Model}
An attacker's goal is to minimise the fair-share bandwidth allocation that a server can give to hosts, and they are expected to act rationally in its pursuit.
Threat actors are external and act intentionally, aren't expected to be 
\emph{advanced persistent threats}, and likely range from hacktivists to moderately funded adversaries.
We assume that attacks will be volumetric DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a transit-link attack).
The addresses of the set of unwitting reflector nodes are visible to the target, though any bots taking part in an attack or the machines those bots control are not revealed to the target without communication with \nth{3} party organisations such as upstream ISPs.
The discovery of any reflector by some defence system does have a cost to the attacker---there is a particularly large (yet finite) supply of viable reflector nodes \cite{DBLP:conf/ndss/Rossow14}, but the constraints that each has a large upstream bandwidth and support for high-amplification-factor protocols narrow this pool.

We do not assume that an attacker has white-box access to an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in volatile network traffic statistics falls outside of the scope of this work.
The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/corr/abs-1902-09062}.
These are APT-level capabilities, whose exploration presents a rich source for future work.

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to design agents such that filtering decisions are computed per flow.
However, these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

\subsection{System Design and Assumptions}
A deployment environment is a network with a set of \emph{ingress/egress points} from its domain of control, through which traffic can flow, and a set of protected \emph{destination nodes}.
These nodes may be services, servers, or in the case of Autonomous Systems (ASes) and transit networks, egress points leading to other networks.
{\color{revisiontext}\cbstart Agents are co-located with each egress switch (i.e., $k$ ingress points from other ASes require $k$ agents), all employ the same action model/design}, and control the proportion of upstream packets from each external host to discard.
Each destination node $s$ has a maximum capacity, \cbend $U_s$.

We assume that the deployment environment is a moderately complex software-defined network, because the paradigm offers features which can directly benefit RL agents acting within.
The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners (e.g., as examined by \textcite{TNSM-that-guys-paper}).
Agents communicate with their co-hosted OpenFlow-enabled switches---running a modified version of \emph{Open vSwitch} (OVS) \cite{open-vswitch}---to install probabilistic packet-drop rules.

\cbstart
{\color{revisiontext}
Our system design applies to both software-defined and traditional networks of arbitrary shape and size.
Only the ingress/egress nodes from a network need to be OpenFlow-enabled, as it is advantageous to perform filtering as close to a flow's source as possible.
In a traditional network, each agent has exclusive control over its switch's tables.
In a fully software-defined network, these agents require exclusive control over the first table, forwarding legitimate packets to subsequent tables managed by the network's controller.
The main difference is that a traditional network needs this additional hardware, and does not allow an operator to dynamically determine where the ``edge'' of their network lies through vNF relocation.}
\cbend

\subsection{Algorithm}
To make decisions cheaply and at low latency, we use \emph{semi-gradient Sarsa with tile coding} as described in \cref{eqn:sg-sarsa} and \cref{sec:reinforcement-learning}, rather than using neural networks or more complicated function approximators.
Exploration is introduced via $\epsilon$-greedy action selection, linearly annealing $\epsilon$ to 0 over time.
Each agent has its own internal parameter vector $\wvec{}$, and agents do not share their weight vector updates with one another (but may share experience and traces with one another).

Although the choice of a classical RL method likely brings lower theoretical performance, there are significant reasons to favour such methods; these include lower latency decision-making, lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
This aligns with our goal of quick online learning, and faster adaptation to aggregate changes in traffic without introducing dedicated tensor processing hardware to networks.
Simpler decision boundaries reduce the risk of overfitting and unexpected behaviour, and we expect that the simplicity of tile-based policy computation will also greatly aid interpretability of anomalous action choices.

@@

%?? How is global state acquired? See \cref{fig:global-state-path}. Why take paths in the way that we do? Mention deterministic ECMP routing etc...
Global state is a vector of load values in $\mathbb{R}^4$ (\si{\mega\bit\per\second}) depending upon the bandwidth measurements regularly received from monitors in the environment.
For any flow, an agent then computes the path it would take through the network.
The incoming load recorded along the first hop, last hop, and tertiles of the path may then be tile-coded together.
\cbstart {\color{revisiontext}In the event that the path from an agent to its destination is shorter than 4 hops, we duplicate (in order of preference) the load measurement of a middle hop or the last hop.} \cbend
\Cref{fig:global-state-path} illustrates the process.

@@

We describe and analyse each of the per-flow features included in the state vector throughout \cref{sec:rethinking-the-state-space}.
Each feature is tiled separately, with the exception of packet in/out count (per-window and total), mean in/out packet size, and $\Delta$ in/out rate, which are combined with the last action taken.
\cbstart {\color{revisiontext}Rather than having the network push the data to an agent, the agent requests this information about active flows periodically to isolate it from non-control-plane traffic and to eliminate the risk of resource exhaustion by excessive requests.} \cbend

@@

\subsection{Action Space}
When monitoring a source-destination pair, an agent uses its state vector to decide which proportion of that flow's \emph{inbound} traffic should be dropped.
%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
This is implemented by installing an action via OpenFlow, instructing its host switch to drop each relevant packet with probability $p$.
We choose to drop packets rather than impose traffic limits as it offers us a discrete action space without prior knowledge of traffic characteristics or measurement.
Furthermore, we need not consider burstiness, fairness or tuning (such as per-flow bucket sizes) which could limit scalability.
We offer two models on how to choose $p$:

\subsubsection{Instant control}
Each agent directly chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
These choices ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
Since this model needs no forward planning, we found it best to set the discount factor $\gamma=0$ (making agents purely myopic).

\subsubsection{Guarded control}
The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$.
The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and congestion-unaware traffic as required.

@@

\subsection{Systems Considerations}\label{sec:systems-considerations}
Taking many actions per timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
This introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated.
At their worst, these risks present additional attack surface to an adversary.
To adapt to these problems, we make use of \emph{timed random sequential} updates.

Each agent begins with an empty work list.
For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
Uncompleted work is passed on to the next timestep, until the list is emptied, at which point it is repopulated using the set of available measurements.
To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields.
Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 3 flows in this time.
We expect this should be tuned based on the frequency at which statistics arrive.
\cbstart
{\color{revisiontext}Naturally, this implies that an agent must carry work forward (and coalesce state updates) when \emph{host density} is $n>3$ (\cref{sec:evaluation}); this behaviour is not explicitly a property of network size.}
\cbend

@@

\begin{figure}
	\centering
	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-cap-box}
	\vspace{-0.75cm}
	\caption{
		Learned performance of Instant Control agents when benign traffic is TCP-like, using only a single feature as a basis for decisions.
		All of the chosen features can offer a marked improvement over no protection at all.
		Global state and Mean IAT still offer the greatest improvement above baseline, but packet-level statistics are considerably less effective for this class of traffic.
		\label{fig:tcp-cap-feature-plots}
	}
\vspace{-0.7cm}
\end{figure}

\begin{figure}
	\centering
%	\vspace{-0.25cm}
	\includegraphics[width=0.65\linewidth]{../plots/ftprep-tcp-laf-cap-box}
	\vspace{-0.75cm}
	\caption{
		Learned performance of Instant Control agents when benign traffic is TCP-like, combining each feature with the last action taken as a basis for decisions.
		This combination causes a significant improvement in the effectiveness of packet-level and per-window statistics.
		\label{fig:tcp-laf-feature-plots}
	}
\vspace{-0.6cm}
\end{figure}

The main element required by a per-source model is a feature set with high predictive power, so that behavioural differences are apparent to an agent.
Elaborating on the statistics discussed in \cref{sec:motivation} which others have shown to be effective, we believe the following features to be useful (and humanly justifiable), and investigate their use alongside different traffic types:
%?? We use these features, and why...

@@

\fakepara{$\mathbf{\Delta}$ Send/receive rate}
The change in traffic rates caused by the last action.
Behavioural changes induced by bandwidth expansion/reduction are expected to be most visible here.

@@

\fakepara{Mean packet size per window}
Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other flooding attacks.
